### 2023-09-18 paper/paper.qmd
Cross referenced first figure

---
title: "Elephant Seal Age JAE Intraspecific Variation"
format: html
editor: visual
---

```{r}
#| label: setup

library(tidyverse)
library(ggthemes)
sealdat <- read_csv(here::here("data/raw/128L pull 2023_06_29.csv"), 
                    show_col_types = FALSE) %>% 
  mutate(observed = if_else(observed == "B", "Breeder", "Non-breeder"))

```

## Introduction

## Methods

## Results

Pup sex shifted towards males as maternal age increased (@fig-repro_age).

```{r}
#| label: fig-repro_age
#| fig-cap: "Pup sex shifted towards males with increasing maternal age."

sealdat %>% 
  filter(pupsex %in% c("M", "F"), 
         age < 18) %>% 
  group_by(age, pupsex) %>%
  count(pupsex) %>% 
  pivot_wider(values_from = n, names_from = pupsex) %>% 
  mutate(num = M + F,
         percf = F / num,
         se = (percf * (1 - percf) / num)^0.5) %>% 
  ggplot(aes(x = age, y = percf, weight = num)) +
  geom_point() +
  geom_errorbar(aes(ymin = percf - se, ymax = percf + se), width = 0.2) +
  geom_hline(yintercept = 0.5, lty = 2) +
  stat_smooth(method = "glm", method.args = list(family = "binomial"), formula = y ~ x) +
  geom_text(aes(label = num, vjust = 1, y = 0.03)) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  labs(x = "Maternal Age (Years)", y = "% Pups Female") +
  theme_few()

```

## Discussion

##