### 2023-09-19 paper/paper.qmd
Added two more reproductive success figures

---
title: "Elephant Seal Age JAE Intraspecific Variation"
format: 
  html: 
    echo: false 
editor: visual
bibliography: references.bib
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(ggthemes)
library(gridExtra)

sealdat <- read_csv(here::here("data/raw/128L pull 2023_06_29.csv"), 
                    show_col_types = FALSE) %>% 
  mutate(observed = if_else(observed == "B", "Breeder", "Non-breeder"))

```

## Introduction

This paper is basically the same as another paper about polar bears (*Ursus maritimus*) [@naciri2022].

## Methods

## Results

### Changes to annual cycle timing with increased maternal age

In order to create the timing figure, I need a csv with these columns: breeding arrival DOY, breeding departure DOY, molt arrival DOY, molt departure DOY.

-   breeding arrival DOY = firstobsbreeddoy

-   breeding departure DOY = firstobsbreeddoy + breeddur

-   *molt arrival DOY = get from Roxanne*

-   molt departure DOY = lastobsmoltdoy

### Reproductive success and maternal age

As maternal age increases, pup sex skews slightly towards males. Pup survival and recruitment are not significantly affected by maternal age (@fig-repro_age).

```{r}
#| label: fig-repro_age
#| caption: "Pup sex skews slightly towards males with increasing maternal age. Pup survival decreases slightly, but survival and recruitment are not significantly affected by maternal age.

#Pup sex shifts towards males with increasing maternal age.

pup_sex <- sealdat %>% 
  filter(pupsex %in% c("M", "F"), 
         age < 18) %>% 
  group_by(age, pupsex) %>%
  count(pupsex) %>% 
  pivot_wider(values_from = n, names_from = pupsex) %>% 
  mutate(num = M + F,
         percf = F / num,
         se = (percf * (1 - percf) / num)^0.5) %>% 
  ggplot(aes(x = age, y = percf, weight = num)) +
  geom_point() +
  geom_errorbar(aes(ymin = percf - se, ymax = percf + se), width = 0.2) +
  geom_hline(yintercept = 0.5, lty = 2) +
  stat_smooth(method = "glm", method.args = list(family = "binomial"), formula = y ~ x) +
  geom_text(aes(label = num, vjust = 1, y = 0.03)) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  labs(x = "Maternal Age (Years)", y = "% Pups Female") +
  theme_few()

#Pup survival doesn't change with increased maternal age.

pup_survival <- sealdat %>% mutate(pup_surv = ifelse(pupseeneveragain > 0, 1, 0)) %>% 
  drop_na(pup_surv) %>% 
  group_by(age, pup_surv) %>%
  count(pup_surv) %>% 
  pivot_wider(names_from = pup_surv, values_from = n) %>% 
  mutate(n = (`0`+`1`),
         perc = `0` / n,
         se = (perc * (1-perc) / n)^0.5) %>% 
  filter(n > 5) %>% 
  ggplot(aes(x = age, y = perc)) +
  geom_point() +
  geom_smooth(method = "glm", method.args = list(family = binomial), formula = y ~ x) +
  geom_errorbar(aes(ymin = perc - se, ymax = perc + se), width = 0.2) +
  geom_text(aes(label = n, 
                vjust = -1), y = .5) +
  labs(x = "Maternal Age (Years)", y = "Fraction Pups Not Seen Again") +
  theme_few()

#code that I haven't got to work yet that shows this by sex
#ggplot(aes(x = age, y = perc, weight = n, col = pupsex)) +
  # geom_text(pupsex == "M", aes(label = n, 
  #                              vjust = -1, 
  #                              col = pupsex), y = 0.5) + 
  # geom_text(pupsex == "F", aes(label = n, 
  #                              vjust = -1, 
  #                              col = pupsex), y = 0.5) +

#Pup recruitment doesn't change with maternal age

pup_recruit <- sealdat %>% 
  filter(pupsex == "F") %>% 
  group_by(age, puprecruited) %>%
  drop_na(puprecruited) %>% 
  count(puprecruited) %>% 
  pivot_wider(names_from = puprecruited, values_from = n) %>% 
  mutate(n = (`0` + `1`),
         perc = `0` / n,
         se = (perc * (1 - perc) / n)^0.5) %>% 
  filter(n > 5) %>% 
  ggplot(aes(x = age, y = perc, weight = n)) +
  geom_point() + 
  xlim(2.5 , 13.5) +
  geom_errorbar(aes(ymin = perc - se, ymax = perc + se), width = 0.2) +
  stat_smooth(method = "glm", method.args = list(family = "binomial"), formula = y ~ x) +
  labs(x = "Maternal Age (Years)", y = "Fraction Pups Never Breed") +
  theme_few() +
  geom_text(aes(label = n,
                vjust = 1,
                y = 1))

grid.arrange(pup_sex, pup_survival, pup_recruit)

```

## Discussion

##