### 2023-09-20 paper/paper.qmd
Updating and making the colors consistent throughout the document

---
title: "Elephant Seal Age JAE Intraspecific Variation"
format: 
  html: 
    echo: false 
editor: visual
bibliography: references.bib
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(ggthemes)
library(cowplot)

sealdat <- read_csv(here::here("data/raw/128L pull 2023_06_29.csv"), 
                    show_col_types = FALSE) %>% 
  mutate(observed = if_else(observed == "B", "Breeder", "Non-breeder"))

```

## Introduction

Our hypotheses:

1.  Elephant seal reproductive success declines with age.

2.  Reproductive declines persist to the next generation (maternal effect senescence).

3.  Seals optimize their annual cycle timing to compensate for reproductive declines.

Note: This paper is similar to another paper about polar bears (*Ursus maritimus*) [@naciri2022].

## Methods

## Results

### H1: Age distribution and breeding success

```{r}
#| label: fig-age_dist
#| fig-cap: "A: Age structure of northern elephant seals at Año Nuevo. B: Percentage of each age class that were observed breeding."
#| fig-width: 6
#| fig-height: 8

#Age distribution panel A

age_dist <- sealdat %>% 
  group_by(age) %>% 
  count(age) %>% 
  ggplot(aes(x = age, y = n)) + 
  geom_col(fill = "#829BD1") + 
  labs(x = "Number of individuals", y = "Age (Years)") + 
  theme_few()

#Panel B: percentage of each age class that were breeders or  nonbreeders

breed_nonbreed <- sealdat %>% 
  group_by(age) %>% 
  summarize(n_observed = sum(!is.na(observed)), 
            perc_breed = sum(observed == "Breeder") / n_observed, 
            se = (perc_breed * (1 - perc_breed) / n_observed)^0.5) %>% 
  ggplot(aes(x = age, y = perc_breed)) + 
  stat_smooth(method = "lm", 
              formula = y ~ poly(x , 2), 
              se = TRUE, 
              color = "#849ED5", 
              fill = "#C8D3ED") +
  geom_point() + 
  geom_errorbar(aes(ymin = perc_breed - se, ymax = perc_breed + se), width = 0) +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) + 
  labs(x = "Maternal Age (Years)", y = "% Breeders") + 
  theme_few()

plot_grid(age_dist, breed_nonbreed, ncol = 1, labels = "AUTO")

```

### H2: Maternal effect senescence

As maternal age increases, pup sex skews slightly towards males. Pup survival declines slightly with maternal age, but pup recruitment is not significantly affected by maternal age (@fig-repro_age). Reproductive declines therefore do not persist to the next generation.

```{r}
#| label: fig-repro_age
#| fig-cap: "Pup sex skews slightly towards males with increasing maternal age. Pup survival decreases slightly, but survival and recruitment are not significantly affected by maternal age."
#| fig-width: 6
#| fig-height: 8

#A: Pup sex shifts towards males with increasing maternal age.

pup_sex <- sealdat %>% 
  filter(pupsex %in% c("M", "F"), 
         age < 18) %>% 
  group_by(age, pupsex) %>%
  count(pupsex) %>% 
  pivot_wider(values_from = n, names_from = pupsex) %>% 
  mutate(num = M + F,
         percf = F / num,
         se = (percf * (1 - percf) / num)^0.5) %>% 
  ggplot(aes(x = age, y = percf, weight = num)) +
  geom_hline(yintercept = 0.5, lty = 2) +
  stat_smooth(method = "glm", 
              method.args = list(family = "binomial"), 
              formula = y ~ x, 
              color = "#829BD1", 
              fill = "#C8D3ED") +
  geom_point() +
  geom_errorbar(aes(ymin = percf - se, ymax = percf + se), width = 0) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  labs(x = "Maternal Age (Years)", y = "% Pups Female") +
  theme_few()

#B Pup survival and recruitment doesn't change with increased maternal age.

pup_surv_recruit <- sealdat %>% 
  mutate(pup_surv = ifelse(pupseeneveragain > 0, 1, 0)) %>% 
  filter(pupsex == "F") %>% 
  group_by(age) %>%
  summarize(n_surv = sum(!is.na(pup_surv)),
            mean_surv = sum(pup_surv, na.rm = TRUE) / n_surv,  
            se_surv = (mean_surv * (1 - mean_surv) / n_surv)^0.5,
            n_recruit = sum(!is.na(puprecruited)), 
            mean_recruit = sum(puprecruited, na.rm = TRUE) / n_recruit, 
            se_recruit = (mean_recruit * (1 - mean_recruit) / n_recruit)^0.5)  %>%  
  pivot_longer(cols = c(mean_surv, se_surv, mean_recruit, se_recruit), 
               names_to = c("stat", "metric"), 
               names_pattern = "(.*)_(.*)") %>% 
  pivot_wider(names_from = "stat", 
              values_from = "value") %>% 
  mutate(low = mean - se, 
         high = mean + se, 
         age = age + ifelse(metric == "surv", -0.1, 0.1), 
         metric = factor(metric, 
                         levels = c("recruit", "surv"), 
                         labels = c("Recruitment", "1st Year Survival"))) %>% 
  filter(n_surv > 5) %>% 
  ggplot(aes(x = age, y = mean, color = metric)) + 
  geom_smooth(aes(weight = n_surv), 
              method = "glm", 
              method.args = list(family = binomial), 
              formula = y ~ x, 
              fill = "#C8D3ED") +
  geom_pointrange(aes(ymin = low, ymax = high)) +
  scale_color_manual(values = c("#829BD1", "#E9B668")) + 
  scale_y_continuous(labels = scales::percent, limits = c(0, .5)) + 
  labs(x = "Maternal Age (Years)", y = "% Pups") +
  theme_few() + 
  theme(legend.position = c(0.1,0.97), 
        legend.justification = c(0,1), 
        legend.title = element_blank(), 
        legend.direction = "horizontal")

plot_grid(pup_sex, pup_surv_recruit, ncol = 1, labels = "AUTO")

```

### H3: Changes to annual cycle timing with increased maternal age

In order to create the timing figure, I need a csv with these columns: breeding arrival DOY, breeding departure DOY, molt arrival DOY, molt departure DOY.

-   breeding arrival DOY = firstobsbreeddoy

-   breeding departure DOY = firstobsbreeddoy + breeddur

-   molt arrival DOY = lastobsmoltdoy - moltdur

-   molt departure DOY = lastobsmoltdoy

```{r}
#| label: annual_timing

```

## Discussion

##