### 2023-11-25 paper/paperv2.qmd
Figure updates

---
title: "Elephant Seal Age JAE Intraspecific Variation"
format: 
  html: 
    echo: false
    fig-width: 4.5
    fig-height: 4.5
    code-fold: true
    code-tools: true
    warning: false
editor: visual
bibliography: references.bib
---

```{r}
#| label: setup
#| include: false

library(ggthemes)
library(cowplot)
library(lme4)
library(lmerTest) # Loading this library will automatically append p-values to lmer models estimated using something called Satterthwaite's method
library(merTools)
library(tidyverse)
library(ggeffects)
library(patchwork)

sealdat <- read_csv(here::here("data/raw/128L pull 2023_11_21.csv"), 
                    show_col_types = FALSE) %>% 
  mutate(observed = if_else(observed == "B", "Breeder", "Non-breeder"), 
         observed_int = if_else(observed == "Breeder", 1, 0), 
         pup_survived = pupseeneveragain > 0) %>% 
  filter(age > 3)

```

# Introduction

Our hypotheses:

1.  Elephant seal reproductive success declines with age.

2.  Reproductive declines persist to the next generation (maternal effect senescence).

3.  Seal phenology is shifted later for older animals and breeders.

**Note**: *This paper is similar to another paper about polar bears (*Ursus maritimus*)* [@naciri2022].

# Results

### H1: Age distribution and breeding status

We predicted that breeding success would decline with age. We found that breeding success increases to prime age, then decreases at older ages (@fig-age_dist).

```{r}
#| label: age_breeding_mod

breeding_mod_bad <- glmer(observed_int ~ poly(age, 2) + (1 | animalID) + (1 | year),
                      data = filter(sealdat), 
                      family = "binomial")

count(sealdat, year) %>% 
     left_join(as_tibble(ranef(breeding_mod_bad)$year, 
                         rownames = "year") %>% 
                   mutate(year = as.numeric(year)) %>% 
                   rename(coef = `(Intercept)`),
               by = "year") %>%
     ggplot(aes(n, coef, color = year)) +
     geom_point() +
     scale_color_gradient(low = "blue", high = "red") +
     theme_few()

breeding_mod <- glmer(observed_int ~ poly(age, 2) + (1 | animalID) + (1 | year),
                      data = filter(sealdat, year >= 1990), 
                      family = "binomial")
breed_pred <- ggpredict(breeding_mod, terms = c("age [all]"))
summary(breeding_mod)

```

```{r}
#| label: fig-age_dist
#| fig-cap: "A: Age structure of northern elephant seals at Año Nuevo. B: Percentage of each age class that were observed breeding."
#| fig-width: 9

#Age distribution panel A

age_dist <- sealdat %>% 
  filter(age < 22,
         year >= 1990) %>% 
  group_by(age) %>% 
  count(age) %>% 
  ggplot(aes(x = age, y = n)) + 
  geom_col(fill = "#829BD1") + 
  labs(x = "Age (Years)", y = "Number of individuals") + 
  theme_few() + 
  theme(axis.title = element_text(size = 13))

#Panel B: percentage of each age class that were breeders or  nonbreeders

breed_nonbreed <- sealdat %>% 
  filter(year >= 1990) %>% 
  group_by(age) %>% 
  summarize(n_observed = sum(!is.na(observed)), 
            perc_breed = sum(observed == "Breeder") / n_observed, 
            se = (perc_breed * (1 - perc_breed) / n_observed)^0.5) %>% 
  filter(age < 22) %>%
  ggplot(aes(x = age, y = perc_breed)) + 
  geom_ribbon(aes(x, ymin = conf.low, ymax = conf.high),
              breed_pred,
              inherit.aes = FALSE,
              fill = "#829BD1",
              alpha = 0.25) +
  geom_line(aes(x, predicted),
            breed_pred,
            inherit.aes = FALSE,
            color = "#829BD1",
            linewidth = 1) +
  geom_point() + 
  geom_errorbar(aes(ymin = perc_breed - se, ymax = perc_breed + se), width = 0) +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
  labs(x = "Maternal Age (Years)", y = "% Breeders (Mean ± SE)") + 
  theme_few() + 
  theme(axis.title = element_text(size = 13))

plot_grid(age_dist, breed_nonbreed, nrow = 1, labels = "AUTO")

ggsave("outputs/agefig1.png", plot = , height = 4, width = 9)

```

Population level prediction - the fact that the raw data falls above it comes from the fact that those age classes are correlated with years and individuals that were very successful.

```{r}
#| label: fig-breeding_bias

year_coef <- as_tibble(ranef(breeding_mod)$year, 
                       rownames = "year") %>% 
    mutate(year = as.numeric(year)) %>% 
    rename(coef = `(Intercept)`)

animal_coef <- as_tibble(ranef(breeding_mod)$animalID, 
                         rownames = "animalID") %>% 
    mutate(animalID = as.numeric(animalID)) %>% 
    rename(coef = `(Intercept)`)

year_bias <- sealdat %>% 
  left_join(year_coef, by = "year") %>% 
  mutate(age = factor(age)) %>% 
  ggplot(aes(age, coef)) +
  geom_boxplot() +
  geom_hline(yintercept = 0, color = "#829BD1", linetype = "dashed") +
  scale_x_discrete("Maternal Age (Years)", breaks = seq(5, 20, by = 5)) +
  ylab("Random effect of year") +
  theme_few()

animalID_bias <- sealdat %>% 
  left_join(animal_coef, by = "animalID") %>% 
  mutate(age = factor(age)) %>% 
  ggplot(aes(age, coef)) +
  geom_boxplot() +
  geom_hline(yintercept = 0, color = "#829BD1", linetype = "dashed") +
  scale_x_discrete("Maternal Age (Years)", breaks = seq(5, 20, by = 5)) +  
  ylab("Random effect of individual") +
  theme_few()

plot_grid(breed_nonbreed, year_bias, animalID_bias, nrow = 1)

```

### H2: Maternal effect senescence

We predicted that reproductive declines would carry over to the next generation - that is, pups born to older mothers would have decreased survival and recruitment success.

We found that neither pup survival nor recruitment is significantly affected by maternal age (@fig-repro_age).

We reject our hypothesis: reproductive declines do not persist to the next generation, and there is no evidence of maternal effect senescence.

```{r}
#| label: fig-repro_age
#| fig-cap: "Pup survival decreases slightly, but survival and recruitment are not significantly affected by maternal age."
#| fig-width: 9

#B Pup survival and recruitment doesn't change with increased maternal age.

pup_surv_recruit <- sealdat %>% 
  mutate(pup_surv = ifelse(pupseeneveragain > 0, 1, 0)) %>% 
  filter(pupsex == "F", 
         pupyearborn < 2016) %>% 
  group_by(age) %>%
  summarize(n_surv = sum(!is.na(pup_surv)),
            mean_surv = sum(pup_surv, na.rm = TRUE) / n_surv,  
            se_surv = (mean_surv * (1 - mean_surv) / n_surv)^0.5,
            n_recruit = sum(!is.na(puprecruited)), 
            mean_recruit = sum(puprecruited, na.rm = TRUE) / n_recruit, 
            se_recruit = (mean_recruit * (1 - mean_recruit) / n_recruit)^0.5)  %>%  
  pivot_longer(cols = c(mean_surv, se_surv, mean_recruit, se_recruit), 
               names_to = c("stat", "metric"), 
               names_pattern = "(.*)_(.*)") %>% 
  pivot_wider(names_from = "stat", 
              values_from = "value") %>% 
  mutate(low = mean - se, 
         high = mean + se, 
         age = age + ifelse(metric == "surv", -0.1, 0.1), 
         metric = factor(metric, 
                         levels = c("recruit", "surv"), 
                         labels = c("Recruitment", "Survival"))) %>% 
  filter(n_surv > 5) %>% 
  ggplot(aes(x = age, y = mean, color = metric)) + 
  geom_pointrange(aes(ymin = low, ymax = high)) +
  scale_color_manual(values = c("#829BD1", "#E9B668")) + 
  scale_y_continuous(labels = scales::percent, limits = c(0, .5)) + 
  labs(x = "Maternal Age (Years)", y = "% Pups (Mean ± SE)") +
  theme_few() + 
  theme(legend.position = c(0.1,0.97), 
        legend.justification = c(0,1), 
        legend.title = element_blank(), 
        legend.direction = "horizontal", 
        axis.title = element_text(size = 13)
)

pup_surv_recruit
```

Results of the Pup Survival model, with individual and year as random effects:

```{r}
#Panel B Survival Model
surv_dat <- sealdat %>% filter(pupyearborn < 2016)
pup_survival_model <- glmer(pup_survived ~ age + (1 | animalID) + (1 | year),
                            data = surv_dat,
                            family = "binomial")
summary(pup_survival_model)

```

Results of the Pup Recruitment model, with individual and year as random effects:

```{r}
#Panel B Recruitment Model

pup_recruit_model <- glmer(puprecruited ~ age + (1 | animalID) + (1 | year),
                           data = filter(sealdat, year > 1984 & year < 2019), 
                           family = "binomial")
summary(pup_recruit_model)

```

Modeling the above metrics as quadratic, just in case:

```{r}
surv_dat <- sealdat %>% filter(pupyearborn < 2016)
pup_survival_model <- glmer(pup_survived ~ poly(age, 2) + (1 | animalID) + (1 | year),
                            data = surv_dat,
                            family = "binomial")
summary(pup_survival_model)

pup_recruit_model <- glmer(puprecruited ~ poly(age, 2) + (1 | animalID) + (1 | year),
                           data = filter(sealdat, year > 1984 & year < 2019), 
                           family = "binomial")
summary(pup_recruit_model)

```

Neither were significant as 2nd order polynomials.

The model of Survival \~ Maternal age was significant before we treated Year as a random effect. Was that because pups born earlier in the study had more time to be seen again (i.e., a sampling artifact?) If so, the plot of the random effect of year should have a negative slope.

We also plotted survival bias to show that 95% of pups were resighted within 7 years. Both plots are supplemental.

```{r}
#| label: fig-surv_bias
#| fig-cap: "A: Plotting the intercept of the random effect of year shows that variation in ocean conditions is likely driving patterns in survival, rather than age of mother. B: 95% of seals that survived were resighted within ~seven years of their birth."
#| fig-width: 9

year_ranef <- ranef(pup_survival_model)$year %>% 
  as_tibble(rownames = "year") %>% 
  mutate(year = as.numeric(year)) %>% 
  rename(intercept = `(Intercept)`)

year_ranef_plot <- ggplot(year_ranef, aes(year, intercept)) + 
  geom_point() + 
  geom_smooth(method = "loess", 
              formula = y ~ x,
              color = "red") + 
  theme_classic()

survival_bias <- sealdat %>% 
  group_by(animalID) %>% 
  summarize(first_resight_yrs = min(age))

survival95 <- quantile(survival_bias$first_resight_yrs, 0.95)
#n for 95% quantile
sum(survival_bias$first_resight_yrs > 7)

survival_bias_plot <- ggplot(survival_bias, aes(x = first_resight_yrs)) +
  stat_ecdf() +
  geom_vline(xintercept = survival95, linetype = "dashed", color = "red") +
  scale_x_continuous(limits = c(0, 10)) + 
  theme_classic()

plot_grid(year_ranef_plot, survival_bias_plot, nrow = 1, labels = "AUTO")

```

There is a drop off in recent years that seems to support the idea that recent years have a sampling bias, but the earlier portion suggests that overall the variation is caused by ocean conditions and not sampling. See the Cohort paper!

### H3a: Changes in pup sex ratio with increased maternal age

We found that as maternal age increases, pup sex skews slightly towards males - but it's not significant.

```{r}
#| label: fig-sex_ratio
#| fig-cap: "Pup sex skews slightly towards males with increasing maternal age, but it's not significant."
#| fig-width: 9
#A: Pup sex shifts towards males with increasing maternal age, but it's not significant.

pup_sex <- sealdat %>% 
  filter(pupsex %in% c("M", "F"), 
         age < 19) %>% 
  group_by(age, pupsex) %>%
  count(pupsex) %>% 
  pivot_wider(values_from = n, names_from = pupsex) %>% 
  mutate(num = M + F,
         percf = F / num,
         se = (percf * (1 - percf) / num)^0.5) %>% 
  ggplot(aes(x = age, y = percf, weight = num)) +
  geom_hline(yintercept = 0.5, lty = 2, color = "#829BD1") +
  geom_point() +
  geom_errorbar(aes(ymin = percf - se, ymax = percf + se), width = 0) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  labs(x = "Maternal Age (Years)", y = "% Female Pups (Mean ± SE)") +
  theme_few() + 
  theme(axis.title = element_text(size = 13))

pup_sex
```

Results of the model for pup sex ratio:

```{r}
pup_sex_mod <- glmer(as.factor(pupsex) ~ age + (1 | animalID) + (1 | year),
                     data = filter(sealdat, 
                                   pupsex %in% c("M", "F"), 
                                   age < 19), 
                     family = "binomial")
summary(pup_sex_mod)
```

### H3b: Changes to annual cycle timing with increased maternal age

We predicted that seals would shift the timing of their annual cycle.

We found that as maternal age increases, there are slight adjustments to the timing of life history events. These events (breeding/molting arrival/departure) are typically shifted later for older animals (@fig-phenology).

```{r}

seal_phenology <- sealdat %>% 
  mutate(BA = firstobsbreeddoy, 
         BD = firstobsbreeddoy + breeddur, 
         MA = lastobsmoltdoy - moltdur, 
         MD = lastobsmoltdoy) %>% 
  drop_na(age, BA, BD, MA, MD)

phenology_n <- unique(seal_phenology$animalID)

timing_summary <- seal_phenology %>% 
  group_by(age, observed) %>% 
  summarize(across(BA:MD, mean), 
            .groups = "drop") %>% 
  pivot_longer(cols = -c(age, observed),  
               names_to = "event", 
               values_to = "doy") %>% 
  group_by(age, observed) %>% 
  group_modify(\(data, keys) {
    tibble(event = c("Long trip", "Breeding", "Short trip", "Molting", "longtrip2"),
           doy = c(data$doy, 365) - c(0, data$doy))
  }) %>% 
  ungroup() %>% 
  mutate(event = factor(event, levels = c("Long trip", 
                                          "Breeding", 
                                          "Short trip", 
                                          "Molting", 
                                          "longtrip2")))
timing_summary_flipped <- timing_summary %>% 
  mutate(observed = ifelse(observed == "Breeder",
                           "Non-breeder",
                           "Breeder")) %>% 
  filter(event != "longtrip2")

```

The table is supplemental.

```{r}
#| label: fig-phenology
#| fig-height: 7
#| fig-width: 7
#| fig-cap: "Successful breeding and increased maternal age both impact elephant seal annual cycles. A: Visualization of life history timing across ages, faceted by breeding status. White dashed lines on each panel represent the raw data from the other panel to facilitate comparisons between breeders and non-breeders. B: LMM coefficients for phenological parameters. The units for maternal age coefficients are in days per year. The coefficients for breeding status are in days. Black points are significant, gray points are not significant."

phen_plot1 <- ggplot(timing_summary, aes(x = age, y = doy, fill = event)) + 
  geom_area(position = position_stack(reverse = TRUE),
            color = NA) +
  geom_line(position = position_stack(reverse = TRUE),
            color = "white") +
  geom_line(data = timing_summary_flipped,
            position = position_stack(reverse = TRUE),
            color = "white",
            linetype = "dashed",
            alpha = 0.5) +
  scale_fill_discrete(type = c("#829BD1", "#E9B668", "#7EC6CC", "#E091B9", "#829BD1"),
                      breaks = c("Long trip", "Breeding", "Short trip", "Molting")) + 
  scale_x_continuous("Maternal Age (Years)", expand = expansion(mult = 0)) +
  scale_y_continuous("Day of Year", expand = expansion(mult = 0)) +
  facet_grid(cols = vars(observed)) +
  theme_few() + 
  theme(axis.title = element_text(size = 12),
        panel.spacing = unit(1, "lines"))

# Utility functions for model fitting and coefficient extraction
fit_model <- function(dat) {
  lmer(doy ~ age + observed + (1 | animalID) + (1 | year), data = dat)
}
extract_estimate <- function(models, param) {
  map_dbl(models, \(mod) summary(mod)$coefficients[param, "Estimate"])
}
extract_stderr <- function(models, param) {
  map_dbl(models, \(mod) summary(mod)$coefficients[param, "Std. Error"])
}
extract_pval <- function(models, param) {
  map_dbl(models, \(mod) summary(mod)$coefficients[param, "Pr(>|t|)"])
}

# This data frame contains a model for each phenological variable w.r.t. age
phenology_models <- seal_phenology %>% 
  pivot_longer(BA:MD, names_to = "phenology", values_to = "doy") %>% 
  select(phenology, observed, doy, age, animalID, year) %>% 
  mutate(observed = factor(observed, levels = c("Non-breeder", "Breeder"))) %>% 
  nest(model_data = c(doy, age, observed, animalID, year)) %>% 
  mutate(model = map(model_data, fit_model),
         age_coef = extract_estimate(model, "age"),
         age_stderr = extract_stderr(model, "age"),
         age_pval = extract_pval(model, "age"),
         age_signif = age_pval * 8 < 0.05,
         breeder_coef = extract_estimate(model, "observedBreeder"),
         breeder_stderr = extract_stderr(model, "observedBreeder"),
         breeder_pval = extract_pval(model, "observedBreeder"),
         breeder_signif = breeder_pval * 8 < 0.05,)

phen_wrapped <- tibble(
  event_begin = c(-Inf, 1:4),
  event_end = c(1:4, Inf),
  event = timing_summary$event[0:5]
)

phen_plot2 <- phenology_models %>% 
  select(-c(model_data, model)) %>% 
  pivot_longer(c(ends_with("coef"), ends_with("stderr"), ends_with("signif")),
               names_to = c("param", "metric"),
               names_sep = "_",
               values_to = "value") %>% 
  pivot_wider(names_from = metric, values_from = value) %>% 
  mutate(signif = as.logical(signif), 
         param = factor(param, 
                        labels = c("'Delay with maternal age (days year'^-1 * ')'", 
                                   "'Delay with breeding (days)'")), 
         phenology = factor(phenology, 
                            labels = c("Breeding\narrrival", 
                                       "Breeding\ndeparture", 
                                       "Molt\narrrival", 
                                       "Molt\ndeparture"))) %>% 
  ggplot(aes(x = coef, y = phenology)) +
  # Prevent geom_rect() from establishing a continuous y-axis scale
  geom_point(alpha = 0) +
  geom_rect(aes(xmin = -Inf, xmax = Inf,
                ymin = event_begin, ymax = event_end,
                fill = event),
            phen_wrapped,
            inherit.aes = FALSE) +
  geom_pointrange(aes(xmin = coef - stderr, 
                      xmax = coef + stderr,
                      color = signif)) +
  scale_color_manual(values = c("FALSE" = "grey50", "TRUE" = "black"),
                     guide = "none") +
  scale_fill_discrete(type = c("#829BD1", "#E9B668", "#7EC6CC", "#E091B9", "#829BD1"),
                      breaks = c("Long trip", "Breeding", "Short trip", "Molting")) + 
  expand_limits(x = 0) +
  facet_grid(cols = vars(param), 
             scales = "free_x", 
             switch = "x",
             labeller = label_parsed) +
  theme_few() + 
  theme(axis.title = element_blank(),
        legend.direction = "horizontal",
        legend.title = element_blank(),
        strip.placement = "outside",
        strip.text = element_text(size = 12, color = "black"),
        panel.spacing = unit(1, "lines"))

plot_grid(
  get_legend(phen_plot2),
  phen_plot1 + theme(legend.position = "none"),
  phen_plot2 + theme(legend.position = "none"),
  align = "v",
  axis = "lr",
  labels = c("", "A", "B"),
  ncol = 1,
  rel_heights = c(0.1, 1, 1)
)

ggsave("outputs/agefig4.png", plot = , height = 9, width = 9)

```

Supplemental table

```{r}

phenology_models %>% 
  select(-c(model_data, model)) %>% 
  pivot_longer(age_coef:breeder_signif, 
               names_to = c("param", "metric"),
               names_sep = "_",
               values_to = "value") %>% 
  pivot_wider(names_from = metric, values_from = value) %>% 
  mutate(coef = str_glue("{round(coef, 2)} +- {round(stderr, 2)}"),
         signif = str_glue("p = {round(pval, 4)} ({signif == 1})")) %>% 
  select(-c(stderr, pval)) %>% 
  pivot_wider(names_from = param, values_from = c(coef, signif)) %>% 
  select(phenology, ends_with("age"), ends_with("breeder")) %>% 
  knitr::kable(col.names = c("Phenology", 
                             "Age coef. (days/year)", 
                             "Significance", 
                             "Breeding status coef. (days)", 
                             "Significance")) %>% 
  kableExtra::kable_styling()

```