### 2024-07-19 paper/paperv3.qmd
Updating phenology with the breakpoint model

---
title: "Reproduction frequency and offspring survival decline in elephant seals past prime age"
format: 
  html: 
    echo: false
    fig-width: 7
    fig-height: 5
    code-fold: true
    code-tools: true
    warning: false
    toc: true
editor: visual
bibliography: references.bib
---

```{r}
#| label: setup
#| include: false  
library(ggthemes)
library(cowplot)
library(lme4)
library(lmerTest) # Loading this library will automatically append p-values to lmer models estimated using something called Satterthwaite's method
library(merTools)
library(tidyverse)
library(ggeffects)
library(ggridges)
library(patchwork)
source(here::here("R/format.R"))
theme_set(theme_few(base_size = 13))

age_senesce <- 11

sealdat <- read_csv(here::here("data/raw/128L pull 2023_12_05.csv"), 
                    show_col_types = FALSE) %>% 
  mutate(observed = if_else(observed == "B", "Breeder", "Non-breeder"), 
         observed_int = if_else(observed == "Breeder", 1, 0), 
         pup_survived = pupseeneveragain > 0) %>% 
  filter(age > 3, 
         year > 1987) %>% 
  mutate(animalID = factor(animalID),
         year_fct = factor(year),
         age10 = (age - age_senesce) / 10,
         age_cat = factor(age >= age_senesce, 
                          labels = c("Young", "Old"))) %>% 
  group_by(animalID) %>% 
  mutate(firstrepro = ifelse(max(year) < 2020 && any(observed_int == 1),
                             min(age[observed_int == 1]),
                             NA),
         lastrepro = ifelse(max(year) < 2020 && any(observed_int == 1),
                            max(age[observed_int == 1]),
                            NA),
         lastobs = ifelse(max(year) < 2020, max(age), NA)) %>% 
  ungroup()

resight <- read.csv(here::here("data/raw/fullresights.csv")) %>% 
  select(animalID, resightID, season, date, timeofyear, calyear, yday) %>% 
  mutate(animalID = as.factor(animalID))

```

# Introduction

Our hypotheses:

1.  Elephant seal reproductive success declines with age.

2.  Reproductive declines persist to the next generation (maternal effect senescence). That is, offspring survival and reproduction decrease with maternal age.

3.  Offspring sex ratios shift towards males with old age.

4.  Phenology (breeding haulout and molt haulout) is shifted later for older animals and breeders.

# Results

n = `r n_distinct(sealdat$animalID)` distinct animals, with `r n_distinct(sealdat)` total observations.

### Raw data

```{r}
#| label: fig-raw-data
#| fig-width: 9
#| fig-cap: "A: Raw longitudinal data for 1,202 known-age female elephant seals and their observations from birth, to recruitment, to death. B: Histogram of the number of seals in each age class in the dataset."

sighting_img <- sealdat %>% 
  group_by(animalID) %>% 
  summarize(yearborn = yearborn[1],
            first_obs = min(year),
            last_obs = max(year),
            .groups = "keep") %>% 
  reframe(yearborn = yearborn,
          year = yearborn:last_obs,
          age = year - yearborn,
          prerecruit = year < first_obs) %>% 
  left_join(sealdat %>% 
              group_by(animalID, year) %>% 
              summarize(observed_int = max(observed_int),
                        .groups = "drop"),
            by = c("animalID", "year")) %>% 
  mutate(
    observed = case_when(
      age == 0 ~ "Born",
      prerecruit ~ "Pre-recruit",
      observed_int == 0 ~ "Non-breeder",
      observed_int == 1 ~ "Breeder",
      is.na(observed_int) ~ "Unobserved"
    ),
    observed = factor(observed, levels = c("Born",
                                           "Pre-recruit",
                                           "Non-breeder",
                                           "Breeder",
                                           "Unobserved"))
  ) %>% 
  group_by(animalID) %>% 
  mutate(lastrepro = max(age),
         yearborn = na.omit(yearborn)[1]) %>% 
  ungroup() %>% 
  arrange(yearborn, lastrepro, animalID) %>% 
  mutate(animalID = factor(animalID, levels = unique(animalID))) %>% 
  ggplot(aes(year, animalID)) +
  geom_raster(aes(fill = observed)) +
  scale_fill_viridis_d(direction = -1) +
  guides(fill = guide_legend(position = "inside", 
                             direction = "horizontal",
                             ncol = 1,
                             byrow = TRUE)) +
  labs(y ="Individual") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.x = element_blank(),
        legend.title = element_blank(),
        # legend.position.inside = c(0.01, 0.99),
        legend.justification = c(0, 1),
        legend.background = element_blank(),
        legend.key.size = unit(4, "mm"),
        legend.text = element_text(size = 8))

age_dist <- sealdat %>% 
  filter(age < 22) %>% 
  group_by(age) %>% 
  count(age) %>% 
  ggplot(aes(x = age, y = n)) + 
  geom_col(fill = "grey50") + 
  labs(x = "Age (Years)", y = "Number of individuals") + 
  theme()

(sighting_img | age_dist) + plot_layout(widths = c(2, 1)) + plot_annotation(tag_levels = "A")

```

### H1: Reproductive senescence

We predicted that breeding success would decline with age. Based on the literature, we chose age `r age_senesce` to be the end of prime age and indicate the onset of senescence. If breeding success declined with age, we would expect to see a negative slope post-senescence (@fig-breeding-mod).

```{r}
#| label: breeding-mod

# Fit GLMMs to the data with binary variable indicating before and after onset # of senescence. We a priori chose age 11 for senescence (age_senesce).
breed_age_mod <- glmer(
  observed_int ~ age10 : age_cat + (1 | animalID) + (1 | year_fct),
  sealdat,
  family = "binomial",
  control = glmerControl(optimizer = "bobyqa")
)

summary(breed_age_mod)$coefficients

```

```{r}
#| label: fig-breeding-mod
#| fig-cap: "Breeding rates increased until age 10, then decreased from age 11 on. Black points represent mean and standard error of the observed breeding rates. The solid lines and shaded areas represent the mean and 95% confidence interval of the population-level model predictions. The observed data are unequally distributed across years, so the dotted lines indicate the predicted breeding rates averaged over years (i.e., including the random effect of year), weighted by the number of seals observed in each year."

# Summarize the raw data
breed_nonbreed <- sealdat %>%
  group_by(age) %>%
  summarize(n_observed = n(),
            n_breed = sum(observed_int),
            perc_breed = n_breed / n_observed,
            breed_lwr = binom.test(n_breed, n_observed)$conf.int[1],
            breed_upr = binom.test(n_breed, n_observed)$conf.int[2]) %>%
  filter(age <= 20)

#Create sample size data frame
n_breed <- sealdat %>% 
  filter(age <= 20) %>% 
  count(age)

# Create ANNUAL (i.e. rand effect of year only) predictions
breed_pred_years <- ggpredict(
  breed_age_mod, 
  terms = c("age10 [all]", "age_cat", "year_fct"),
  type = "random"
) %>% 
  as_tibble() %>% 
  mutate(age = x * 10 + age_senesce,
         age_cat = factor(group, levels = c("Young", "Old")),
         year_fct = factor(facet)) %>% 
  filter((age_cat == "Young" & age < age_senesce) |
           (age_cat == "Old" & age >= age_senesce),
         age <= 20)

# Create WEIGHTED-AVERAGE predictions using years' random effects
# ggpredict can't do that for confidence intervals
# create prediction grid
breed_grid <- expand_grid(
  age = 4:20,
  year_fct = factor(1988:2023)
) %>% 
  mutate(age10 = (age - age_senesce) / 10,
         age_cat = factor(ifelse(age < age_senesce, "Young", "Old"),
                          levels = c("Young", "Old")))

seal_years <- count(sealdat, year_fct)
breed_pred_wgt <- breed_grid %>% 
  mutate(predicted = predict(breed_age_mod, 
                             newdata = breed_grid,
                             re.form = ~ (1 | year_fct)),
         predicted_se = suppressWarnings(
           # warning: "se.fit computation uses an approximation to estimate the sampling distribution of the parameters"
           predict(breed_age_mod, 
                   newdata = breed_grid,
                   re.form = NA,
                   se.fit = TRUE)
         )$se.fit,
         predicted_pop = predict(breed_age_mod, 
                                 newdata = breed_grid,
                                 re.form = NA),
         conf_lo = predicted - 1.96 * predicted_se,
         conf_hi = predicted + 1.96 * predicted_se) %>% 
  left_join(seal_years, by = "year_fct") %>% 
  group_by(age, age_cat) %>% 
  summarize(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
                   \(x) weighted.mean(x, n)),
            .groups = "drop") %>% 
  # invert logit
  mutate(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
                family(breed_age_mod)$linkinv)) %>% 
  left_join(n_breed, by = "age")

ggplot(breed_pred_wgt, aes(age, predicted)) +
  # Predictions for individual years
  geom_line(aes(group = interaction(age_cat, year_fct)),
            breed_pred_years,
            alpha = 0.1) +
  # Weighted model (CI and mean response)
  geom_ribbon(aes(fill = age_cat, ymin = conf_lo, ymax = conf_hi), 
              alpha = 0.5) +
  geom_line(aes(color = age_cat),
            linewidth = 1.2) +
  # Population-level mean response
  geom_line(aes(y = predicted_pop, group = age_cat),
            linetype = "dotted",
            linewidth = 1.2,
            alpha = 0.5) +
  # Summarized raw data
  geom_pointrange(aes(y = perc_breed,
                      ymin = breed_lwr,
                      ymax = breed_upr),
                  breed_nonbreed) +
  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
  scale_y_continuous("Breeding probability", 
                     labels = scales::percent,
                     limits = c(NA, 1.02),
                     breaks = seq(0.4, 1.0, by = 0.2)) +
  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
  labs(x = "Age (years)") +
  theme(legend.position = "none") + 
  geom_text(aes(label = n), y = 1, vjust = -0.5)

ggsave("outputs/repro_senesce.svg", width = 13, height = 6, units = "in")
```

```{r}
#| label: h1-pred-deets

#Overall decline 11-19
tibble(age = c(11, 19),
       age10 = (age - age_senesce) / 10,
       age_cat = "Old") %>% 
  # logit-scale predictions
  mutate(predicted = predict(breed_age_mod, 
                             newdata = .,
                             re.form = NA),
         se = suppressWarnings(predict(breed_age_mod, 
                                       newdata = .,
                                       se.fit = TRUE,
                                       re.form = NA)$se.fit),
         pred_lwr = predicted - 1.96 * se,
         pred_upr = predicted + 1.96 * se,
         # Invert the logit
         across(c(predicted, pred_upr, pred_lwr), 
                family(breed_age_mod)$linkinv))

#Difference between years 
best_worst_yr <- ranef(breed_age_mod)$year_fct %>% 
  as_tibble(rownames = "year_fct") %>% 
  arrange(`(Intercept)`) %>% 
  slice(1, nrow(.))

tibble(age = c(11, 19),
       age10 = (age - age_senesce) / 10,
       age_cat = "Old") %>% 
  cross_join(best_worst_yr) %>% 
  mutate(predicted = predict(breed_age_mod, 
                             newdata = .,
                             type = "response", 
                             re.form = ~ (1 | year_fct)))
```

Sample size n = `r nrow(sealdat)` observations of `r n_distinct(sealdat$animalID)` individuals.

### H2: Maternal effect senescence

We predicted that reproductive declines would carry over to the next generation - that is, pups born to older mothers would have decreased survival (for female and male pups) and reproductive success (for female pups). Pup survival significantly decreased with maternal age, but pup reproduction did not (@fig-repro-age).

```{r}
#| label: repro-age

# Isolate pup survival/recruitment 
surv_recr_data <- sealdat %>% 
  drop_na(pupID) %>% 
  # Only pups born 2016 or earlier (otherwise survival biased)
  filter(pupyearborn <= 2016) %>% 
  filter(age < 20) %>% 
  mutate(pup_surv_int = as.integer(pup_survived),
         pup_recr_int = as.integer(puprecruited))

# Fit pup *survival* model
pup_surv_mod <- glmer(
  pup_surv_int ~ age10 : age_cat + (1 | animalID) + (1 | year_fct),
  surv_recr_data,
  family = "binomial",
  control = glmerControl(optimizer = "bobyqa")
)

# Fit pup *recruitment* model (female offspring only)
pup_recr_mod <- glmer(
  # NOTE: no random effect for individual bc singular fit
  # In survival model, 93 individuals (15%) had >3 pups (male+female pups)
  # In recruit model, only 17 individuals (4%) had >3 pups (female pups only)
  pup_recr_int ~ age10 : age_cat + (1 | year_fct),
  filter(surv_recr_data, pupsex == "F"),
  family = "binomial",
  control = glmerControl(optimizer = "bobyqa")
)

"Survival"
summary(pup_surv_mod)$coefficients
"Recruitment"
summary(pup_recr_mod)$coefficients

```

```{r}
#| label: fig-repro-age
#| fig-width: 8
#| fig-height: 8

# Summarize the raw data
surv_summ <- surv_recr_data %>% 
  group_by(age) %>% 
  filter(age <= 19) %>% 
  summarize(n_pups = n(), 
            n_surv = sum(pup_surv_int),
            perc_surv = n_surv / n_pups, 
            surv_lwr = binom.test(n_surv, n_pups)$conf.int[1],
            surv_upr = binom.test(n_surv, n_pups)$conf.int[2])
recr_summ <- surv_recr_data %>% 
  filter(pupsex == "F") %>% 
  group_by(age) %>% 
  summarize(n_fpups = n(), 
            n_recr = sum(pup_recr_int),
            perc_recr = n_recr / n_fpups, 
            recr_lwr = binom.test(n_recr, n_fpups)$conf.int[1],
            recr_upr = binom.test(n_recr, n_fpups)$conf.int[2])

#Create sample size data frame
n_surv <- surv_recr_data %>% 
  filter(age <= 19) %>% 
  count(age)

# Create pup survival predictions (population-level and weighted by year)
surv_grid <- expand_grid(
  age = 4:19,
  year_fct = factor(1988:2016)
) %>% 
  mutate(age10 = (age - age_senesce) / 10,
         age_cat = factor(ifelse(age < age_senesce, "Young", "Old"),
                          levels = c("Young", "Old")))

surv_years <- count(surv_recr_data, year_fct)
surv_pred_wgt <- surv_grid %>% 
  mutate(predicted = predict(pup_surv_mod, 
                             newdata = surv_grid,
                             re.form = ~ (1 | year_fct)),
         predicted_se = suppressWarnings(
           # warning: "se.fit computation uses an approximation to estimate the sampling distribution of the parameters"
           predict(pup_surv_mod, 
                   newdata = surv_grid,
                   re.form = NA,
                   se.fit = TRUE),
         )$se.fit,
         predicted_pop = predict(pup_surv_mod, 
                                 newdata = surv_grid,
                                 re.form = NA),
         conf_lo = predicted - 1.96 * predicted_se,
         conf_hi = predicted + 1.96 * predicted_se) %>% 
  left_join(surv_years, by = "year_fct") %>% 
  group_by(age, age_cat) %>% 
  summarize(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
                   \(x) weighted.mean(x, n)),
            .groups = "drop") %>% 
  # invert logit
  mutate(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
                family(pup_surv_mod)$linkinv)) %>% 
      left_join(n_surv, by = "age")

# Predictions for each year (random effect)
surv_pred_years <- ggpredict(
  pup_surv_mod,
  terms = c("age10 [all]", "age_cat", "year_fct [all]"),
  type = "random",
  interval = "confidence"
) %>% 
  as_tibble() %>% 
  mutate(age = x * 10 + age_senesce,
         age_cat = group,
         year_fct = facet) %>% 
  filter((age_cat == "Young" & age < age_senesce) |
           (age_cat == "Old" & age >= age_senesce),
         age <= 20)

fig_surv <- ggplot(surv_pred_wgt, aes(age, predicted)) +
  # Individual years
  geom_line(aes(group = interaction(age_cat, year_fct)),
            surv_pred_years,
            alpha = 0.1) +
  # Weighted model (confidence interval and mean response)
  geom_ribbon(aes(fill = age_cat, ymin = conf_lo, ymax = conf_hi), 
              alpha = 0.2) +
  geom_line(aes(color = age_cat, linetype = age_cat), linewidth = 1.2) +
  # Population-level mean response
  geom_line(aes(y = predicted_pop, group = age_cat),
            linetype = "dotted", linewidth = 1.2, alpha = 0.5) +
  # Raw data
  geom_pointrange(aes(y = perc_surv, 
                      ymin = surv_lwr, 
                      ymax = surv_upr),
                  surv_summ) +
  # Senescence threshold
  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
  scale_y_continuous("Offspring survival", 
                     labels = scales::percent,
                     limits = c(0, 1)) +
  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
  scale_linetype_manual(values = c("dashed", "solid")) +
  labs(x = "Age (years)") +
  theme(legend.position = "none") +
  geom_text(aes(label = n), y = 1, vjust = 1)

#Create sample size data frame
n_recruit <- surv_recr_data %>% 
  filter(pupsex == "F") %>% 
  count(age)

# Create pup recruitment predictions (population-level and weighted by year)
recr_grid <- expand_grid(
  age = 4:19,
  year_fct = factor(1988:2016)
) %>% 
  mutate(age10 = (age - age_senesce) / 10,
         age_cat = factor(ifelse(age < age_senesce, "Young", "Old"),
                          levels = c("Young", "Old")))

recr_years <- surv_recr_data %>% 
  filter(pupsex == "F") %>% 
  count(year_fct)
recr_pred <- recr_grid %>% 
  mutate(predicted = predict(pup_recr_mod, 
                             newdata = recr_grid,
                             re.form = ~ (1 | year_fct)),
         predicted_se = suppressWarnings(
           # warning: "se.fit computation uses an approximation to estimate the sampling distribution of the parameters"
           predict(pup_recr_mod, 
                   newdata = recr_grid,
                   re.form = NA,
                   se.fit = TRUE),
         )$se.fit,
         predicted_pop = predict(pup_recr_mod, 
                                 newdata = recr_grid,
                                 re.form = NA),
         conf_lo = predicted - 1.96 * predicted_se,
         conf_hi = predicted + 1.96 * predicted_se) %>% 
  left_join(recr_years, by = "year_fct") %>% 
  group_by(age, age_cat) %>% 
  summarize(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
                   \(x) weighted.mean(x, n)),
            .groups = "drop") %>% 
  # invert logit
  mutate(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
                family(pup_recr_mod)$linkinv)) %>% 
  left_join(n_recruit, by = "age")

# Predictions for each year (random effect)
recr_pred_years <- ggpredict(
  pup_recr_mod,
  terms = c("age10 [all]", "age_cat", "year_fct [all]"),
  type = "random",
  interval = "confidence"
) %>% 
  as_tibble() %>% 
  mutate(age = x * 10 + age_senesce,
         age_cat = group,
         year_fct = facet) %>% 
  filter((age_cat == "Young" & age < age_senesce) |
           (age_cat == "Old" & age >= age_senesce),
         age <= 20)

fig_recr <- ggplot(recr_pred, aes(age, predicted)) +
  # Individual years
  geom_line(aes(group = interaction(age_cat, year_fct)),
            recr_pred_years,
            alpha = 0.1) +
  # Weighted model (confidence interval and mean response)
  geom_ribbon(aes(fill = age_cat, ymin = conf_lo, ymax = conf_hi), 
              alpha = 0.2) +
  geom_line(aes(color = age_cat), linetype = "dashed", linewidth = 1.2) +
  # Population-level mean response
  geom_line(aes(y = predicted_pop, group = age_cat),
            linetype = "dotted", linewidth = 1.2, alpha = 0.5) +
  # Raw data
  geom_pointrange(aes(y = perc_recr, 
                      ymin = recr_lwr, 
                      ymax = recr_upr),
                  recr_summ) +
  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
  scale_y_continuous("Offspring recruitment", 
                     labels = scales::percent,
                     limits = c(0, 1)) +
  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
  labs(x = "Age (years)") +
  theme(legend.position = "none") + 
  geom_text(aes(label = n), y = 1, vjust = 1)

(fig_surv | fig_recr) + plot_annotation(tag_levels = "A") + plot_layout(ncol = 1, nrow = 2)

```

Sample size for offspring survival was n = `r nrow(surv_recr_data)` observations of `r n_distinct(surv_recr_data$animalID)` individuals. Sample size for offspring recruitment was n = `r nrow(n_recruit)` observations of `r n_distinct(n_recruit$animalID)` female individuals.

```{r}
#| label: h2-pred-deets

#Overall decline 11-19
tibble(age = c(11, 19),
       age10 = (age - age_senesce) / 10,
       age_cat = "Old") %>% 
  # logit-scale predictions
  mutate(predicted = predict(pup_surv_mod, 
                             newdata = .,
                             re.form = NA),
         se = suppressWarnings(predict(pup_surv_mod, 
                                       newdata = .,
                                       se.fit = TRUE,
                                       re.form = NA)$se.fit),
         pred_lwr = predicted - 1.96 * se,
         pred_upr = predicted + 1.96 * se,
         # Invert the logit
         across(c(predicted, pred_upr, pred_lwr), 
                family(pup_surv_mod)$linkinv))

#Difference between years 
best_worst_yr_surv <- ranef(pup_surv_mod)$year_fct %>% 
  as_tibble(rownames = "year_fct") %>% 
  arrange(`(Intercept)`) %>% 
  slice(1, nrow(.))

tibble(age = c(11, 19),
       age10 = (age - age_senesce) / 10,
       age_cat = "Old") %>% 
  cross_join(best_worst_yr_surv) %>% 
  mutate(predicted = predict(pup_surv_mod, 
                             newdata = .,
                             type = "response", 
                             re.form = ~ (1 | year_fct)))
```

### H3: Changes in pup sex ratio with increased maternal age

We found that as maternal age increases, pup sex skews slightly towards males, but it is not significant (@fig-sex-ratio).

```{r}
#| label: sex-ratio

pup_sex_data <- sealdat %>% 
  filter(pupsex %in% c("M", "F"), 
         age < 19) %>% 
  mutate(is_male = as.integer(pupsex == "M"))

sexrat_age_mod <- glmer(
  is_male ~ age10 : age_cat + (1 | animalID) + (1 | year_fct),
  pup_sex_data,
  family = "binomial"
)

summary(sexrat_age_mod)$coefficients

```

```{r}
#| label: fig-sex-ratio
#| fig-cap: "Offspring sex ratio skews towards males with increasing maternal age, but the result is not significant (p = 0.6). Black points show the mean and solid black lines show the standard error of the observed sex ratio."

# Summarize the raw data
sexrat_summ <- pup_sex_data %>% 
  group_by(age) %>% 
  summarize(n_observed = n(), 
            n_male = sum(is_male),
            perc_male = n_male / n_observed, 
            male_lwr = binom.test(n_male, n_observed)$conf.int[1], 
            male_upr = binom.test(n_male, n_observed)$conf.int[2]) %>% 
  filter(age <= 20)

#Make sample size data frame
n_sexrat <- pup_sex_data %>% 
  count(age)

# Create offspring sex ratio predictions (population-level and weighted by year)
sexrat_grid <- expand_grid(
  age = 4:18,
  year_fct = factor(1988:2023)
) %>% 
  mutate(age10 = (age - age_senesce) / 10,
         age_cat = factor(ifelse(age < age_senesce, "Young", "Old"),
                          levels = c("Young", "Old")))

sexrat_years <- count(pup_sex_data, year_fct)
sexrat_pred <- sexrat_grid %>% 
  mutate(predicted = predict(sexrat_age_mod, 
                             newdata = sexrat_grid,
                             re.form = ~ (1 | year_fct)),
         predicted_se = suppressWarnings(
           # warning: "se.fit computation uses an approximation to estimate the sampling distribution of the parameters"
           predict(sexrat_age_mod, 
                   newdata = sexrat_grid,
                   re.form = NA,
                   se.fit = TRUE),
         )$se.fit,
         predicted_pop = predict(sexrat_age_mod, 
                                 newdata = sexrat_grid,
                                 re.form = NA),
         conf_lo = predicted - 1.96 * predicted_se,
         conf_hi = predicted + 1.96 * predicted_se) %>% 
  left_join(sexrat_years, by = "year_fct") %>% 
  group_by(age, age_cat) %>% 
  summarize(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
                   \(x) weighted.mean(x, n)),
            .groups = "drop") %>% 
  # invert logit
  mutate(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
                family(sexrat_age_mod)$linkinv)) %>% 
  left_join(n_sexrat, by = "age")

# Predictions for each year (random effect)
sexrat_years <- ggpredict(
  sexrat_age_mod,
  terms = c("age10 [all]", "age_cat", "year_fct [all]"),
  type = "random",
  interval = "confidence"
) %>% 
  as_tibble() %>% 
  mutate(age = x * 10 + age_senesce,
         age_cat = group,
         year_fct = facet) %>% 
  filter((age_cat == "Young" & age < age_senesce) |
           (age_cat == "Old" & age >= age_senesce),
         age <= 20)

ggplot(sexrat_pred, aes(age, predicted)) +
  # Individual years
  geom_line(aes(group = interaction(age_cat, year_fct)),
            sexrat_years,
            alpha = 0.1) +
  # Weighted model (confidence interval and mean response)
  geom_ribbon(aes(fill = age_cat, ymin = conf_lo, ymax = conf_hi), 
              alpha = 0.2) +
  geom_line(aes(color = age_cat, group = age_cat),
            linetype = "dashed", linewidth = 1.2) +
  # Population-level mean response
  geom_line(aes(y = predicted_pop, group = age_cat),
            linetype = "dotted", linewidth = 1.2, alpha = 0.5) +
  # Raw data
  geom_pointrange(aes(y = perc_male, 
                      ymin = male_lwr, 
                      ymax = male_upr),
                  sexrat_summ) +
  geom_vline(xintercept = age_senesce - 0.5, linetype = "dashed") +
  scale_y_continuous("Male offspring", 
                     labels = scales::percent,
                     limits = c(NA, 0.97)) +
  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
  labs(x = "Age (years)") +
  theme(legend.position = "none") +
  geom_text(aes(label = n), y = .97, vjust = 0)

```

Sample size n = `r nrow(pup_sex_data)` observations of `r n_distinct(pup_sex_data$animalID)` individuals.

**Note**: In the above figure, the population-level and weighted-average would lie directly on top of each other, so we've excluded the weighted average. This is because the random effect of year has very little impact on offspring sex ratio, which makes sense biologically. It's easy to imagine scenarios where reproduction would vary year-to-year, but harder to think of a mechanism linking year and offspring sex ratio.

### Phenology

Does the timing of breeding and molting suggest a mechanistic explanation for observed patterns in reproductive and maternal-effect senescence?

We fit a generalized linear mixed-effects model with the duration of a phase (molting or breeding) as the response variable and age, interacting with phase, as the predictor, using year as a random effect. We attempted to fit a model with random effects for both year and individual, but that failed to converge. We assumed the effect of year on phase duration (molting or breeding) was greater than the effect of individual, so we retained year as the random effect.

$$
\begin{align}
d &\sim NB(\mu, k) \\
\mu &= exp(\eta) \\
\eta &= \beta_0 + \beta_p * a + Zy
\end{align}
$$

Where: $d$ is the duration of the phase in days, $NB$ is the negative binomial distribution with location $\mu$ and dispersion $k$, $\eta$ is the linear predictor, $\beta_0$ is the intercept, $\beta_p$ is the coefficient for age ($a$) by phase (i.e., for $p$ in "molting" or "breeding"), $y$ is the year, and $Z$ is the design matrix relating $y$ to $\eta$.

```{r}
#| label: pheno-model

seal_phenology <- sealdat %>% 
  filter(year >= 2011, 
         moltdur >= 7) %>% 
  mutate(BA = firstobsbreeddoy, 
         BD = firstobsbreeddoy + breeddur, 
         MA = lastobsmoltdoy - moltdur, 
         MD = lastobsmoltdoy) %>% 
  drop_na(age, BA, BD, MA, MD)

pheno_model <- seal_phenology %>% 
  transmute(animalID, 
            year_fct, 
            age10 = (age - age_senesce) / 10, 
            age_cat = factor(age >= age_senesce, 
                          labels = c("Young", "Old")),
            moltdur, 
            breeddur) %>% 
  pivot_longer(moltdur:breeddur, 
               names_to = "phase", 
               values_to = "dur_days") %>% 
  glmer.nb(dur_days ~ age10:phase:age_cat + (1 | year_fct), 
           data = .)

summary(pheno_model)

```

```{r}

#Make sample size data frame
n_phen_age <- count(seal_phenology, age)
n_phen_year <- count(seal_phenology, year_fct)

pheno_grid <- expand_grid(
  age = 4:20,
  year_fct = factor(2011:2023), 
  phase = c("breeddur", "moltdur")
) %>% 
  mutate(age10 = (age - age_senesce) / 10,
         age_cat = factor(ifelse(age < age_senesce, "Young", "Old"),
                          levels = c("Young", "Old")))

pheno_pred <- pheno_grid %>% 
  mutate(predicted = predict(pheno_model, 
                             newdata = pheno_grid,
                             re.form = ~ (1 | year_fct)),
         predicted_se = suppressWarnings(
           # warning: "se.fit computation uses an approximation to estimate the sampling distribution of the parameters"
           predict(pheno_model, 
                   newdata = pheno_grid,
                   re.form = NA,
                   se.fit = TRUE),
         )$se.fit,
         predicted_pop = predict(pheno_model, 
                                 newdata = pheno_grid,
                                 re.form = NA),
         conf_lo = predicted - 1.96 * predicted_se,
         conf_hi = predicted + 1.96 * predicted_se) %>% 
  left_join(n_phen_year, by = "year_fct") %>% 
  group_by(age, age_cat) %>% 
  summarize(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
                   \(x) weighted.mean(x, n)),
            .groups = "drop") %>% 
  # invert logit
  mutate(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
                family(pheno_model)$linkinv)) %>% 
  left_join(n_phen_age, by = "age")

phen_pred_years <- ggpredict(
  pheno_model,
  terms = c("age10 [all]", "age_cat", "phase", "year_fct [all]"),
  type = "random",
  interval = "confidence"
) %>% 
  as_tibble() %>% 
  mutate(age = x * 10 + age_senesce,
         age_cat = group,
         phase = facet,
         year_fct = panel) %>% 
  filter((age_cat == "Young" & age < age_senesce) |
           (age_cat == "Old" & age >= age_senesce))

pheno_raw <- seal_phenology %>% 
  select(breeddur, moltdur, age) %>% 
  pivot_longer(c(breeddur, moltdur), 
               names_to = "phase", 
               values_to = "dur_days") %>% 
  group_by(age, phase) %>% 
  summarize(n = n(),
            dur_days_se = sd(dur_days) / sqrt(n),
            dur_days = mean(dur_days),
            dur_days_lwr = dur_days - dur_days_se,
            dur_days_upr = dur_days + dur_days_se,
            .groups = "drop")

breed_pheno_raw <- pheno_raw[pheno_raw$phase == "breeddur",]
molt_pheno_raw <- pheno_raw[pheno_raw$phase == "moltdur",]

breed_duration_plot <- ggpredict(
  pheno_model, 
  terms = c("age10 [all]", "age_cat", "phase", "year_fct [all]")
) %>% 
  as_tibble() %>%
  transmute(age = x * 10 + age_senesce,
            age_cat = group,
            dur_days = predicted,
            dur_days_lwr = conf.low,
            dur_days_upr = conf.high,
            phase = facet,
            year_fct = panel) %>% 
  left_join(n_phen_year, by = "year_fct") %>% 
  filter(phase == "breeddur",
         (age_cat == "Young" & age < age_senesce) |
           (age_cat == "Old" & age >= age_senesce)) %>% 
  group_by(age, age_cat) %>% 
  summarize(across(starts_with("dur_days"), \(x) weighted.mean(x, n)),
            .groups = "drop") %>% 
  ggplot(aes(age)) +
  # Predictions by year
  geom_line(aes(age, predicted, group = interaction(phase, age_cat, year_fct)), 
            filter(phen_pred_years, phase == "breeddur"),
            alpha = 0.1) + 
  # Weighted model (CI + mean)
  geom_ribbon(aes(ymin = dur_days_lwr, ymax = dur_days_upr, fill = age_cat),
              alpha = 0.25) +
  geom_line(aes(y = dur_days, color = age_cat, linetype = age_cat)) +
  # Summarized data
  geom_pointrange(aes(y = dur_days, 
                      ymin = dur_days_lwr, 
                      ymax = dur_days_upr),
                  breed_pheno_raw,
                  position = position_dodge(width = 0.5)) +
  scale_linetype_manual(values = c("solid", "dashed")) +
  labs(x = "Maternal age (years)",
       y = "Breeding haul out (days)") +
  theme(legend.position = "none",
        axis.text.y = element_text(size = 8))

molt_duration_plot <- ggpredict(
  pheno_model, 
  terms = c("age10 [all]", "age_cat", "phase", "year_fct [all]")
) %>% 
  as_tibble() %>%
  transmute(age = x * 10 + age_senesce,
            age_cat = group,
            dur_days = predicted,
            dur_days_lwr = conf.low,
            dur_days_upr = conf.high,
            phase = facet,
            year_fct = panel) %>% 
  left_join(n_phen_year, by = "year_fct") %>% 
  filter(phase == "moltdur",
         (age_cat == "Young" & age < age_senesce) |
           (age_cat == "Old" & age >= age_senesce)) %>% 
  group_by(age, age_cat) %>% 
  summarize(across(starts_with("dur_days"), \(x) weighted.mean(x, n)),
            .groups = "drop") %>% 
  ggplot(aes(age)) +
  # Predictions by year
  geom_line(aes(age, predicted, group = interaction(phase, age_cat, year_fct)), 
            filter(phen_pred_years, phase == "moltdur"),
            alpha = 0.1) + 
  # Weighted model (CI + mean)
  geom_ribbon(aes(ymin = dur_days_lwr, ymax = dur_days_upr, fill = age_cat),
              alpha = 0.25) +
  geom_line(aes(y = dur_days, color = age_cat, linetype = age_cat)) +
  # Summarized data
  geom_pointrange(aes(y = dur_days, 
                      ymin = dur_days_lwr, 
                      ymax = dur_days_upr),
                  molt_pheno_raw,
                  position = position_dodge(width = 0.5)) +
  scale_linetype_manual(values = c("solid", "dashed")) +
  labs(x = "Maternal age (years)",
       y = "Molt haul out (days)") +
  theme(legend.position = "none",
        axis.text.y = element_text(size = 8))

```

```{r}
#| label: fig-pheno
#| fig-cap: "Molting gets shorter but breeding doesn't. A: Observed annual time allocation averaged over individuals. B: Raw data for breeding and molting durations (mean ± se) and predictions (mean and 95% CI) from negative binomial weighted by number of seals obersved per year."
#| fig-width: 9
#| fig-height: 9

phenology_unique_animals <- unique(seal_phenology$animalID)
n_phenology <- seal_phenology %>% 
  count(year)

phen_dist <- seal_phenology %>% 
  # group_by(age) %>% 
  count(age) %>% 
  ggplot(aes(x = age, y = n)) + 
  geom_col(fill = "grey50") + 
  labs(x = "Maternal age (Years)", y = "Number of individuals") + 
  theme()

timing_summary <- seal_phenology %>% 
  group_by(age) %>% 
  summarize(across(BA:MD, mean)) %>% 
  pivot_longer(cols = -age,  
               names_to = "event", 
               values_to = "doy") %>% 
  group_by(age) %>% 
  group_modify(\(data, keys) {
    tibble(event = c("Foraging", "Breeding haulout", "Post-breeding", "Molt haulout", "foraging2"),
           doy = c(data$doy, 365) - c(0, data$doy))
  }) %>% 
  ungroup() %>% 
  mutate(event = factor(event, levels = c("Foraging", 
                                          "Breeding haulout", 
                                          "Post-breeding", 
                                          "Molt haulout", 
                                          "foraging2"))) %>% 
    left_join(n_phen_age, by = "age")

annual_plot <- ggplot(timing_summary, aes(x = age, y = doy, fill = event)) + 
  geom_area(position = position_stack(reverse = TRUE),
            color = NA) +
  geom_line(position = position_stack(reverse = TRUE),
            color = "white") +
  scale_fill_discrete(
    type = c("#829BD1", "#D05794", "#829BD1", "#E09A2D", "#829BD1"),
    breaks = c("Foraging", "Breeding haulout", "Molt haulout")
  ) + 
  scale_x_continuous("Maternal age (years)", expand = expansion(mult = 0)) +
  scale_y_continuous("Day of year", expand = expansion(mult = 0)) +
    geom_text(aes(label = "Breeding haul out"), 
            x = 8,
            y = 30,
            color = "white") +
    geom_text(aes(label = "Molting haul out"), 
            x = 8,
            y = 135,
            color = "white") + 
      geom_text(aes(label = "Foraging"), 
            x = 8,
            y = 250,
            color = "white") + 
  theme(panel.spacing = unit(1, "lines"), 
        legend.position = "none")

pheno_raw <- seal_phenology %>% 
  select(breeddur, moltdur, age) %>% 
  pivot_longer(c(breeddur, moltdur), 
               names_to = "phase", 
               values_to = "dur_days") %>% 
  group_by(age, phase) %>% 
  summarize(n = n(),
            dur_days_se = sd(dur_days) / sqrt(n),
            dur_days = mean(dur_days),
            dur_days_lwr = dur_days - dur_days_se,
            dur_days_upr = dur_days + dur_days_se,
            .groups = "drop")

(annual_plot | phen_dist | breed_duration_plot | molt_duration_plot) + plot_annotation(tag_levels = "A") + plot_layout(ncol = 2, nrow = 2)
```

Sample size n = `r nrow(seal_phenology)` observations of `r n_distinct(phenology_unique_animals)` individuals.

# Supplemental Material

## Detailed model outputs

### H1: Reproductive senescence

```{r}
#| label: detailed-h1

summary(breed_age_mod)

```

### H2: Maternal effect senescence

```{r}
#| label: detailed-h2

summary(pup_surv_mod)
summary(pup_recr_mod)

```

### H3: Offspring sex ratio

```{r}
#| label: detailed-h3

summary(sexrat_age_mod)

```

## Models' sensitivity to age of senescence

```{r}
#| label: mod-sens

fit_models <- function(a) {
  age_senesce <- a
  sealdat <- sealdat %>% 
    mutate(age_cat = factor(age >= age_senesce, 
                            labels = c("Young", "Old")))

  # H1
  breed_age_mod <- glmer(
    observed_int ~ age10 : age_cat + (1 | animalID) + (1 | year_fct),
    sealdat,
    family = "binomial",
    control = glmerControl(optimizer = "bobyqa")
  )

  # H2
  surv_recr_data <- sealdat %>% 
    drop_na(pupID) %>% 
    filter(pupyearborn <= 2016) %>% 
    mutate(pup_surv_int = as.integer(pup_survived),
           pup_recr_int = as.integer(puprecruited))

  pup_surv_mod <- glmer(
    pup_surv_int ~ age10 : age_cat + (1 | animalID) + (1 | year_fct),
    surv_recr_data,
    family = "binomial",
    control = glmerControl(optimizer = "bobyqa")
  )

  pup_recr_mod <- glmer(
    pup_recr_int ~ age10 : age_cat + (1 | year_fct),
    filter(surv_recr_data, pupsex == "F"),
    family = "binomial",
    control = glmerControl(optimizer = "bobyqa")
  )

  # H3
  pup_sex_data <- sealdat %>% 
    filter(pupsex %in% c("M", "F"), 
           age < 19) %>% 
    mutate(is_male = as.integer(pupsex == "M"))

  sexrat_age_mod <- glmer(
    is_male ~ age10 : age_cat + (1 | animalID) + (1 | year_fct),
    pup_sex_data,
    family = "binomial"
  )

  list(
    h1 = summary(breed_age_mod)$coefficients,
    h2_surv = summary(pup_surv_mod)$coefficients,
    h2_recr = summary(pup_recr_mod)$coefficients,
    h3 = summary(sexrat_age_mod)$coefficients
  )
}

# Fit models with cut-off at ages 8-13
cutoff <- 8:13
mod_sens <- map(cutoff, fit_models)

# Are hypotheses supported by the model?
extract_p <- \(h) map_dbl(mod_sens, \(m) pnorm(m[[h]]["age10:age_catOld", "z value"]))
p_tbl <- tibble(age_cutoff = cutoff, 
                "H1 (breeding)" = extract_p("h1"),
                "H2a (survival)" = extract_p("h2_surv"),
                "H2b (recruitment)" = extract_p("h2_recr"),
                "H3 (sex ratio)" = extract_p("h3"))

knitr::kable(p_tbl)

```

```{r}
#| label: fig-mod-sens
#| fig-cap: Model sensitivity to onset of senescence. Age at onset of senescence on the x-axis, p-value for the coefficient of the interaction (age category = post-senescence) x (response variable) on the y-axis. Statistical significance (p = 0.05) indicated by dashed line. Hypotheses H1 (reproductive senescence), H2b (maternal-effect senescence; offspring recruitment), and H3 (offspring sex ratio) were insensitive to choice of age for onset of senescence. Hypothesis H2a (maternal-effect senescence; offspring survival) was significantly supported for onset of senescence age 11+. At age 10, the p-vale for the coefficient of the interaction (age category = post-senescence) x (offspring survival) was 0.058.

p_tbl %>% 
  pivot_longer(-age_cutoff, names_to = "hypothesis", values_to = "p") %>%
  ggplot(aes(age_cutoff, p)) +
  geom_point() +
  geom_hline(yintercept = 0.05, linetype = "dashed") +
  geom_text(aes(x, y, label = tag),
            tibble(hypothesis = colnames(p_tbl)[-1],
                   tag = LETTERS[1:4],
                   x = 8, y = 3),
            vjust = 1,
            fontface = "bold") +
  scale_y_log10(labels = scales::number) +
  labs(x = "Senescence onset (years)") +
  expand_limits(y = 3) +
  facet_wrap(~hypothesis)

```

## Selective appearance and disappearance

Age-dependent life history shifts can result from within- or between-animal changes [@vandepol2006]. We hypothesized age-dependent changes in reproduction were due to senescence, i.e. within-animal changes. However, it is possible that selective appearance or disappearance of animals in the population can create the same pattern. For example, if animals the breed more frequently have lower survival, then older animals will reproduce less frequently not because of senescence but rather because infrequent breeders disproportionately survive to older ages (i.e. selective disappearance). We tested whether our significant results (reproductive senescence, maternal effect senescence in pup survival) were influenced by selective appearance and disappearance, following the methods of @vandepol2006, by adding terms for age at first and last reproduction, respectively. If our results are due to selective appearance/disappearance instead of senescence, then the 95% CI for the coefficient of age in the old age class in these models will overlap 0.

```{r}
#| label: fig-selective
#| fig-cap: The coefficient for age in the old age class in models accounting for selective appearance and disappearance. Red and blue indicate  reproductive senescence (H1) and maternal effect senescence (H2a), respectively. Compared to the base model, the appearance model includes a term for age at first reproduction, the disappearance model includes a term for age at last reproduction, and the combined model includes both.

h1_selective_models <- map(
  list(
    # Base model
    observed_int ~ age10 : age_cat + (1 | animalID) + (1 | year_fct),
    # Selective appearance
    observed_int ~ age10 : age_cat + firstrepro + 
      (1 | animalID) + (1 | year_fct),
    # Selective disappearance
    observed_int ~ age10 : age_cat + lastrepro + 
      (1 | animalID) + (1 | year_fct),
    # Selective appearance and disappearance
    observed_int ~ age10 : age_cat + firstrepro + lastrepro +
      (1 | animalID) + (1 | year_fct)
  ),
  \(form) {
    glmer(form, 
          drop_na(sealdat, lastrepro), 
          family = "binomial", 
          control = glmerControl(optimizer = "bobyqa"))
  }
)

h2_selective_models <- map(
  list(
    # Base model
    pup_surv_int ~ age10 : age_cat + (1 | animalID) + (1 | year_fct),
    # Selective appearance
    pup_surv_int ~ age10 : age_cat + firstrepro + 
      (1 | animalID) + (1 | year_fct),
    # Selective disappearance
    pup_surv_int ~ age10 : age_cat + lastrepro + 
      (1 | animalID) + (1 | year_fct),
    # Selective appearance and disappearance
    pup_surv_int ~ age10 : age_cat + firstrepro + lastrepro +
      (1 | animalID) + (1 | year_fct)
  ),
  \(form) {
    glmer(form, 
          drop_na(surv_recr_data, lastrepro), 
          family = "binomial", 
          control = glmerControl(optimizer = "bobyqa"))
  }
)

get_old_coef <- function(model) {
  old_coef <- summary(model)$coefficients["age10:age_catOld", "Estimate"]
  old_se <- summary(model)$coefficients["age10:age_catOld", "Std. Error"]
  ci <- old_coef + c(-1.96, 1.96) * old_se
  tibble(mean = old_coef, 
         lwr = ci[1], 
         upr = ci[2])
}

h1_selective_coef <- map(h1_selective_models, get_old_coef) %>% 
  list_rbind() %>% 
  mutate(model = c("Base", "Appearance", "Disappearance", "Combined"),
         hypothesis = "H1")

h2_selective_coef <- map(h2_selective_models, get_old_coef) %>% 
  list_rbind() %>% 
  mutate(model = c("Base", "Appearance", "Disappearance", "Combined"), 
         hypothesis = "H2")

selective_coef <- rbind(h1_selective_coef, h2_selective_coef) %>% 
  mutate(model = factor(model, levels = c("Base", "Appearance", "Disappearance", "Combined")))

ggplot(selective_coef, aes(model, mean, color = hypothesis)) + 
  geom_pointrange(aes(ymin = lwr, ymax = upr), position = position_dodge(width = 0.2)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_color_manual(values = c("firebrick", "cornflowerblue")) +
  expand_limits(y = 0) + 
  labs(y = "Coefficient", x = "Model") + 
  theme(legend.position = "none")

```

There was no evidence for selective appearance or disappearance influencing the relationship between maternal age and probability of reproduction or offspring survival. The 95% CI of the coefficient for age in the old age class did not overlap 0 for either hypothesis in any of the selective (dis)appearance models [@fig-selective]. Therefore the observed patterns are consistent with the senescence hypothesis (i.e., within-animal change with age).

## Chronological versus biological age

```{r}
#| label: bioage

bioage_senesce <- -7
sealdat2 <- sealdat %>% 
  drop_na(lastobs) %>% 
  mutate(bioage = age - lastobs,
         bioage10 = (bioage - bioage_senesce) / 10,
         bioage_cat = factor(bioage < bioage_senesce,
                             labels = c("Old", "Young")))

```

The chronological age of an individual may not reflect their biological age because aging rates vary within populations [@martin2011]. Years-to-death, a proxy for biological age, may therefore be a better predictor for senescence patterns than chronological age. We repeated our analysis for the reproductive senescence hypothesis using biological age, setting the breakpoint for senescence at biological age -7 years (i.e. 7 years prior to final observation), based on visual inspection of the raw data. This required us to limit our sample to animals with known longevity (i.e., last observation in 2020 or earlier), which reduced our sample size to `r nrow(sealdat2)` observations of `r n_distinct(sealdat2$animalID)` distinct animals. Subsequently, we fit the model for breakpoints at all biological ages -12 to -3 years, to assess whether our results were sensitive to the choice of breakpoint. Finally, we re-fit the chronological reproductive senescence model using the known-longevity dataset and compared chronological versus biological age models using AIC.

```{r}
#| label: bioage-mod

# Biological age model
breed_bioage_mod <- glmer(
  observed_int ~ bioage10 : bioage_cat + (1 | animalID) + (1 | year_fct),
  sealdat2,
  family = "binomial",
  control = glmerControl(optimizer = "bobyqa")
)

summary(breed_bioage_mod)$coefficients

bioage_est <- summary(breed_bioage_mod)$coef["bioage10:bioage_catOld", "Estimate"]
bioage_se <- summary(breed_bioage_mod)$coef["bioage10:bioage_catOld", "Std. Error"]
bioage_coef <- sprintf("%0.3f (%0.3f - %0.3f)", 
                       bioage_est,
                       bioage_est - 1.96 * bioage_se,
                       bioage_est + 1.96 * bioage_se)

old_z <- summary(breed_bioage_mod)$coefficients["bioage10:bioage_catOld", "z value"]
bioage_pval <- sprintf("%0.3f", pnorm(-old_z))

# Breakpoint comparison
p_by_tau <- map_dbl(-12:-3, \(tau) {
  mod <- sealdat %>%
    drop_na(lastobs) %>%
    mutate(bioage = age - lastobs,
           bioage10 = (bioage - tau) / 10,
           bioage_cat = factor(bioage < tau,
                               labels = c("Old", "Young"))) %>% 
    glmer(
      observed_int ~ bioage10 : bioage_cat + (1 | animalID) + (1 | year_fct),
      .,
      family = "binomial",
      control = glmerControl(optimizer = "bobyqa")
    )
  old_z <- summary(mod)$coef["bioage10:bioage_catOld", "z value"] 
  pnorm(old_z)
})

tibble(`Biological age threshold` = -12:-3,
       `p value` = p_by_tau)

# Model comparison
breed_age_mod2 <- glmer(
  observed_int ~ age10 : age_cat + (1 | animalID) + (1 | year_fct),
  drop_na(sealdat2, lastobs),
  family = "binomial",
  control = glmerControl(optimizer = "bobyqa")
)

chrono_bio_aic <- AIC(breed_bioage_mod, breed_age_mod2)
daic <- chrono_bio_aic$AIC - min(chrono_bio_aic$AIC)
chrono_bio_rellik <- exp(-0.5 * daic)
chrono_bio_aicw <- chrono_bio_rellik / sum(chrono_bio_rellik)

```

```{r}
#| label: fig-bio-age
#| fig-cap: Breeding probability over biological age. Points and error bars are the mean ± SE of observed breeding proportions within age classes. The vertical dashed line indicates the threshold for senescence. Solid lines and ribbons are the mean and 95% CI of the population-level breeding probability, estimated by a GLMM with random effects for individual and year. The dashed lines represent the mean of yearly breeding probabilities, weighted by the number of seals observed each year, which should match the observed data more closely than the population-level probabilities.

# Summarize the raw data
bioage_summ <- sealdat2 %>% 
  group_by(bioage) %>% 
  summarize(n_obs = n(), 
            n_breed = sum(observed_int),
            perc_breed = n_breed / n_obs, 
            breed_lwr = binom.test(n_breed, n_obs)$conf.int[1], 
            breed_upr = binom.test(n_breed, n_obs)$conf.int[2])

# Create ANNUAL (i.e. rand effect of year only) predictions
bioage_pred_years <- ggpredict(
  breed_bioage_mod, 
  terms = c("bioage10 [all]", "bioage_cat", "year_fct"),
  type = "random"
) %>% 
  as_tibble() %>% 
  mutate(bioage = x * 10 + bioage_senesce,
         bioage_cat = factor(group, levels = c("Young", "Old")),
         year_fct = factor(facet)) %>% 
  filter((bioage_cat == "Young" & bioage < bioage_senesce) |
           (bioage_cat == "Old" & bioage >= bioage_senesce),
         bioage <= 20)

# Create WEIGHTED-AVERAGE predictions using years' random effects
# ggpredict can't do that for confidence intervals
# create prediction grid
bioage_grid <- expand_grid(
  bioage = -17:0,
  year_fct = factor(1988:2019)
) %>% 
  mutate(bioage10 = (bioage - bioage_senesce) / 10,
         bioage_cat = factor(ifelse(bioage < bioage_senesce, "Young", "Old"),
                          levels = c("Young", "Old")))

bioage_years <- count(sealdat2, year_fct)
bioage_pred_wgt <- bioage_grid %>% 
  mutate(predicted = predict(breed_bioage_mod, 
                             newdata = bioage_grid,
                             re.form = ~ (1 | year_fct)),
         predicted_se = suppressWarnings(
           # warning: "se.fit computation uses an approximation to estimate the sampling distribution of the parameters"
           predict(breed_bioage_mod, 
                   newdata = bioage_grid,
                   re.form = NA,
                   se.fit = TRUE)
         )$se.fit,
         predicted_pop = predict(breed_bioage_mod, 
                                 newdata = bioage_grid,
                                 re.form = NA),
         conf_lo = predicted - 1.96 * predicted_se,
         conf_hi = predicted + 1.96 * predicted_se) %>% 
  left_join(bioage_years, by = "year_fct") %>% 
  group_by(bioage, bioage_cat) %>% 
  summarize(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
                   \(x) weighted.mean(x, n)),
            .groups = "drop") %>% 
  # invert logit
  mutate(across(c(predicted, predicted_pop, conf_lo, conf_hi), 
                family(breed_bioage_mod)$linkinv)) %>% 
  left_join(count(sealdat2, bioage), by = "bioage")

ggplot(bioage_pred_wgt, aes(bioage, predicted)) +
  # Predictions for individual years
  geom_line(aes(group = interaction(bioage_cat, year_fct)),
            bioage_pred_years,
            alpha = 0.1) +
  # Weighted model (CI and mean response)
  geom_ribbon(aes(fill = bioage_cat, ymin = conf_lo, ymax = conf_hi), 
              alpha = 0.5) +
  geom_line(aes(color = bioage_cat),
            linewidth = 1.2) +
  # Population-level mean response
  geom_line(aes(y = predicted_pop, group = bioage_cat),
            linetype = "dotted",
            linewidth = 1.2,
            alpha = 0.5) +
  # Summarized raw data
  geom_pointrange(aes(y = perc_breed,
                      ymin = breed_lwr,
                      ymax = breed_upr),
                  bioage_summ) +
  geom_vline(xintercept = bioage_senesce - 0.5, linetype = "dashed") +
  scale_y_continuous("Breeding probability", 
                     labels = scales::percent,
                     limits = c(NA, 1.02),
                     breaks = seq(0.4, 1.0, by = 0.2)) +
  scale_color_manual(values = c("#7fbc41", "#de77ae")) +
  scale_fill_manual(values = c("#7fbc41", "#de77ae")) +
  labs(x = "Biological age (years before death)") +
  theme(legend.position = "none") + 
  geom_text(aes(label = n), y = 1, vjust = -0.5)

```

The biological age model did not support the reproductive senescence hypothesis. The coefficient for biological age among older animals (biological age \> -7 years) was `r bioage_coef` (95% CI) which was not significantly different than 0 (*p* = `r bioage_pval`). Choice of breakpoint for senescence did not affect the significance of this coefficient. For breakpoints -12 to -3 years, *p* reached a minimum at biological age -7 years (see table above).

The chronological age model outperformed the biological age model according to AIC. The difference in AIC between the two models was `r sprintf("%0.1f", -diff(chrono_bio_aic$AIC))` and the chronological age model had an AIC weight of `r sprintf("%0.3f", chrono_bio_aicw[2])`.

Based on these results, we find there is evidence to support the reproductive senescence hypothesis based on chronological age. The biological age proxy, years-to-death, should be a better predictor of senescence than chronological age if most mortality is age-related. However, adult elephant seal mortality is largely influenced by extrinsic factors, primarily food availability [@colegrove2005; @holser2021]. Poor foraging conditions during El Niño years, for example, may increase mortality rates among animals whose biological age is not yet advanced. As result, the years-to-death proxy for biological age likely conflates many biologically young and old animals that died due to extrinsic factors. More accurate measures of biological age, such as telomere length [@jylhävä2017], may be necessary to investigate senescence in elephant seals and other species with high extrinsic sources of mortality from a non-chronological perspective.

## Observation frequency

Not all seals were detected in all years. A non-detection year during the seal's lifespan (i.e. an unobserved year between the first and last observed years) could represent an unobserved animal at Año Nuevo or temporary emigration. In either case, the animal may or may not have reproduced that year. By excluding these unobserved animals, our inferences for H1 (reproductive senescence) may be biased if detectability varies with age and/or by breeding state. Here, we assess those potential biases.

```{r}
#| label: missing-seals

## Are we missing animals more or less with age?
missing_seals <- sealdat %>% 
  # Need to see a seal 3+ years for calculation to make sense
  group_by(animalID) %>% 
  mutate(n_years = n_distinct(year)) %>% 
  ungroup() %>% 
  filter(n_years >= 3,
         age <= 20) %>% 
  # Fill in non-resighted years
  group_by(animalID) %>% 
  arrange(year) %>% 
  reframe(tibble(year2 = (min(year) + 1):(max(year) - 1),
                 unobserved = as.integer(!year2 %in% year),
                 age = year2 - yearborn[1],
                 age10 = (age - age_senesce) / 10)) %>% 
  ungroup() %>% 
  rename(year = year2) %>% 
  mutate(age_cat = factor(age >= age_senesce, 
                          labels = c("Young", "Old")),
         year_fct = factor(year)) %>% 
  left_join(select(sealdat, animalID, year, observed_int),
            by = c("animalID", "year"))

missing_mod <- glmer(unobserved ~ age10 + (1 | year_fct) + (1 | animalID), 
                     missing_seals, 
                     family = binomial)

summary(missing_mod)

missing_age_se <- summary(missing_mod)$coef["age10", "Std. Error"]
missing_age_est <- summary(missing_mod)$coef["age10", "Estimate"]
missing_age_coef <- sprintf("%0.2f (95%% CI: %0.2f - %0.2f)",
                            missing_age_est,
                            missing_age_est - 1.96 * missing_age_se,
                            missing_age_est + 1.96 * missing_age_se)

missing_mod_old <- glmer(
  unobserved ~ age10 + (1 | year_fct),
  filter(missing_seals, age >= 11),
  family = binomial
)

summary(missing_mod_old)

missing_old_se <- summary(missing_mod_old)$coef["age10", "Std. Error"]
missing_old_est <- summary(missing_mod_old)$coef["age10", "Estimate"]
missing_old_coef <- sprintf("%0.2f (95%% CI: %0.2f - %0.2f)",
                            missing_old_est,
                            missing_old_est - 1.96 * missing_old_se,
                            missing_old_est + 1.96 * missing_old_se)

```

First, we tested whether probability of detection was related to age. We found the proportion of unobserved animals increased with age. For this analysis, we excluded the first and last year of observation for each animal, because animals were necessarily observed in those years. For all years between, we determined whether animals were observed. Then we estimated the effect of age on probability of an animal being unobserved using a GLMM (binomial response, logit link) with random effects for year and animal ([@fig-unobserved]). The coefficient for age was `r missing_age_coef` (*p*\<0.0001). The old-animals-only model estimated the age coefficient to be `r missing_old_coef`.

```{r}
#| label: fig-unobserved
#| fig-cap: The probability of an animal being unobserved increased with age. Points and error bars represent the proportion of unobserved animals within age classes plus/minus one standard error. The blue line and shaded area are the mean and 95% CI probability of an animal being unobserved estimated with a generalized linear mixed effect model. See text for model details.

missing_grid <- expand_grid(age = 5:19, 
                            year_fct = factor(1989:2022)) %>% 
  mutate(age10 = (age - age_senesce) / 10)

missing_pred <- missing_grid %>% 
  mutate(predicted = predict(missing_mod, 
                             newdata = .,
                             re.form = ~ (1 | year_fct)),
         se = suppressWarnings(predict(missing_mod, 
                                       newdata = .,
                                       se.fit = TRUE,
                                       re.form = NA)$se.fit),
         conf.low = predicted - 1.96 * se,
         conf.high = predicted + 1.96 * se) %>% 
  left_join(count(missing_seals, year_fct), by = "year_fct") %>% 
  group_by(age) %>% 
  summarize(across(c(predicted, conf.low, conf.high), mean)) %>% 
  mutate(across(c(predicted, conf.low, conf.high), family(missing_mod)$linkinv))

missing_pred_years <- missing_grid %>% 
  mutate(predicted = predict(missing_mod, 
                             newdata = .,
                             type = "response",
                             re.form = ~ (1 | year_fct)))

missing_pred_old <- expand_grid(age = 11:19, 
                                year_fct = factor(1989:2022)) %>% 
  mutate(age10 = (age - age_senesce) / 10) %>% 
  mutate(predicted = predict(missing_mod_old, 
                             newdata = .,
                             re.form = ~ (1 | year_fct)),
         se = suppressWarnings(predict(missing_mod, 
                                       newdata = .,
                                       se.fit = TRUE,
                                       re.form = NA)$se.fit),
         conf.low = predicted - 1.96 * se,
         conf.high = predicted + 1.96 * se) %>% 
  left_join(count(filter(missing_seals, age >= 11), year_fct), 
            by = "year_fct") %>% 
  group_by(age) %>% 
  summarize(across(c(predicted, conf.low, conf.high), mean)) %>% 
  mutate(across(c(predicted, conf.low, conf.high), family(missing_mod)$linkinv))

missing_pred_pop <- tibble(age = 5:19,
                           age10 = (age - age_senesce) / 10) %>% 
  mutate(predicted = predict(missing_mod, 
                             newdata = .,
                             type = "response",
                             re.form = NA))

missing_seals %>% 
  group_by(age) %>% 
  summarize(n = n(),
            n_unobserved = sum(unobserved),
            unobserved_mean = n_unobserved / n,
            unobserved_lwr = binom.test(n_unobserved, n)$conf.int[1],
            unobserved_upr = binom.test(n_unobserved, n)$conf.int[2]) %>% 
  ggplot(aes(age)) +
  geom_line(aes(y = predicted, group = year_fct), 
            missing_pred_years, 
            alpha = 0.2) +
  # Overall model
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
              missing_pred,
              alpha = 0.3,
              fill = "firebrick") +
  geom_line(aes(y = predicted),
            missing_pred,
            color = "firebrick",
            linewidth = 2) +
  # Old only
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
              missing_pred_old,
              alpha = 0.3,
              fill = "cornflowerblue") +
  geom_line(aes(y = predicted),
            missing_pred_old,
            linetype = "dashed",
            color = "cornflowerblue",
            linewidth = 2) +
  # Raw data
  geom_pointrange(aes(y = unobserved_mean, 
                      ymin = unobserved_lwr, ymax = unobserved_upr)) +
  scale_x_continuous("Age", breaks = seq(5, 19, by = 4)) +
  scale_y_continuous("Unobserved proportion", 
                     labels = scales::label_percent(accuracy = 1),
                     limits = c(0, NA))
```

```{r}
#| label: haulout

breed_haulout <- sprintf(
  "%i - %i", 
  quantile(sealdat$breeddur[sealdat$observed_int == 1], 0.25, na.rm = TRUE), 
  quantile(sealdat$breeddur[sealdat$observed_int == 1], 0.75, na.rm = TRUE)
)

non_breed_haulout <- sprintf(
  "%i - %i", 
  quantile(sealdat$breeddur[sealdat$observed_int == 0], 0.25, na.rm = TRUE), 
  quantile(sealdat$breeddur[sealdat$observed_int == 0], 0.75, na.rm = TRUE)
)

resight_by_observed <- resight %>% 
  mutate(animalID = as.character(animalID)) %>% 
  right_join(select(sealdat, animalID, year, observed_int), 
             by = c("animalID", season = "year"),
             relationship = "many-to-many") %>% 
  count(animalID, season, observed_int) %>% 
  group_by(observed_int) %>% 
  summarize(n_q1 = quantile(n, 0.25), 
            n_mean = mean(n), 
            n_q3 = quantile(n, 0.75))

breed_resight <- with(resight_by_observed[2, ],
                      sprintf("%0.1f (%i - %i)", n_mean, n_q1, n_q3))
nonbreed_resight <- with(resight_by_observed[1, ],
                         sprintf("%0.1f (%i - %i)", n_mean, n_q1, n_q3))

```

The pattern of unobserved animals increasing with age means it is highly unlikely our inferences about reproductive senescence are due to partial detection. If there was no reproductive senescence in the population, then the apparent senescence we observed would have to result from older breeders being less detectable. However, the biology of northern elephant seals makes breeding a more detectable behavior than non-breeding. Animals must come to land for an extended period to give birth and wean their pups. Among the animals in this study, breeders hauled out for longer than non-breeders (1st - 3rd quartile: breeders `r breed_haulout` days; non-breeders `r non_breed_haulout` days). This extended period of time on the beach makes them more accessible to observers. Furthermore, northern elephant seals exhibit a high degree of site fidelity. Once they recruit to a breeding colony, temporary emigration is rare (cite me). The only plausible explanation for unobserved animals being disproportionately breeders would be crowding at the breeding colonies, which could obscure identifying marks. However, breeding animals were resighted `r breed_resight` times per year, compared to `r nonbreed_resight` for non-breeders (mean, 1st - 3rd quartile), which suggests that breeders are more readily observable than non-breeders. Therefore, it is biologically implausible that the unobserved seals were disproportionately breeders and our estimates for reproductive senescence are, if anything, conservative.

## Breakpoint vs quadratic

### H1: Reproductive senescence

```{r}
#| label: quad-h1

breed_age_quad <- glmer(
  observed_int ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year_fct),
  sealdat,
  family = "binomial",
  control = glmerControl(optimizer = "bobyqa")
)

h1_age10_peak <- -summary(breed_age_quad)$coefficients[2, 1] / 
  (2 * summary(breed_age_quad)$coefficients[3, 1])
h1_age_peak <- h1_age10_peak * 10 + age_senesce

summary(breed_age_quad)$coefficients

sprintf("Age peak (H1) = %0.1f", h1_age_peak)

```

### H2: Maternal effect senescence

```{r}
#| label: quad-h2

# H2a
surv_age_quad <- glmer(
  pup_surv_int ~ poly(age10, 2, raw = TRUE) + (1 | animalID) + (1 | year_fct),
  surv_recr_data,
  family = "binomial",
  control = glmerControl(optimizer = "bobyqa")
)

h2a_age10_peak <- -summary(surv_age_quad)$coefficients[2, 1] / 
  (2 * summary(surv_age_quad)$coefficients[3, 1])
h2a_age_peak <- h2a_age10_peak * 10 + age_senesce

summary(surv_age_quad)$coefficients

sprintf("Age peak (H2a) = %0.1f", h2a_age_peak)

# H2b

recr_age_quad <- glmer(
  pup_recr_int ~ poly(age10, 2, raw = TRUE) + (1 | year_fct),
  filter(surv_recr_data, pupsex == "F"),
  family = "binomial",
  control = glmerControl(optimizer = "bobyqa")
)

h2b_age10_peak <- -summary(recr_age_quad)$coefficients[2, 1] / 
  (2 * summary(recr_age_quad)$coefficients[3, 1])
h2b_age_peak <- h2b_age10_peak * 10 + age_senesce

summary(recr_age_quad)$coefficients

sprintf("Age peak (H2b) = %0.1f", h2b_age_peak)

```